Finite Complete Suites for CSP Refinement Testing

### 2018-11-16 scp-journal/article/introduction.tex
revised sections 1 and 2, shortenned paper

\subsection{Motivation}

Model-based testing (MBT) is an active research field; results are currently
being evaluated and integrated into industrial verification processes by many
companies worldwide. 
This holds particularly for the embedded and
cyber-physical systems domains, where critical systems require rigorous
testing~\cite{jp2018ets,DBLP:conf/isola/0001BH18}.

@@

Generation techniques for complete test suites have been developed for
various formalisms, we mention here representative work in the domains of 
finite state machines~\cite{hierons_testing_2004,simao_reducing_2012}, timed automata~\cite{Springintveld2001}, Circus~\cite{DBLP:journals/acta/CavalcantiG11}, 
Timed CSP~\cite{Schneider:1995:OST:203471.203475},
general labelled transition systems~\cite{DBLP:journals/cn/Tretmans96}, symbolic state machines~\cite{DBLP:conf/icst/Petrenko16}, and Kripke structures~\cite{Huang2017}. In this article, we tackle \emph{Communicating Sequential
Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
process algebra that has been shown to be well-suited for the description of
reactive control systems in many publications over almost five decades. Many
of these applications are described in~\cite{Roscoe2010} and in the
references there. Industrial success has also been reported; see, for
example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.

@@

\subsection{Main Contributions}

This article presents   complete black-box test suites for software and systems
modelled using CSP. They can be generated for non-terminating,
divergence-free, finite-state CSP processes with finite alphabets,
interpreted both in the trace and the failures semantics. Divergence freedom
is usually assumed in black-box testing, since it cannot distinguish between
divergence and deadlock.
%
The  novel main contributions in this article may be summarised as follows.
\begin{enumerate}
\item It is shown that trace or failures conformance can be established with finitely many test cases, provided that suitable fault domains are chosen, such that the true behaviour of the SUT is reflected by members of these domains.

@@

\subsection{Related Work}

Our results complement and extend work previously published  
in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}. None of these previous results provide
sufficient conditions 
for constructing finite complete test suites. Consequently, they also do not provide complexity bounds on the number of test executions needed to establish conformance between SUT and reference process.

In~\cite{DBLP:conf/pts/CavalcantiS17}, 
also fault domains are used, but these contain all processes refining a ``top-level'' fault domain process. This concept is orthogonal to the one investigated in the present article:
members of our fault domain need not be in refinement relation to any other process in the domain, they just adhere to the same upper bound $q$ of nodes in their normalised transition graphs.

@@

In~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}, the authors devised linear test cases: after running through a preset trace $s$, test cases for trace refinement check for illegal acceptance of a specific event $e$, and test cases verifying  nondeterministic behaviour check for  acceptance of events from some set $A$. In the present article, we follow the alternative approach proposed in~\cite{DBLP:conf/fm/PeleskaS96,peleska1997a} and use {\it adaptive} test cases. This means that each test case
adapts its trace execution to the nondeterministic behaviour of the SUT, checks for 
  trace violations at any point during the test execution, and checks for the acceptance 
of a given minimal hitting set of events after any legal trace of a test case-specific length.
The adaptive test cases have the advantage that test executions only lead to an 
inconclusive result if the reference process allows for a nondeterministic choice between deadlock and trace continuation in a certain state. In contrast to this, the linear test cases 
may lead to many more futile executions with inconclusive results, if the SUT refuses 
to engage into the next event $e$ 
from the preset trace $s$, due to legal nondeterministic choices leading to a refusal of $e$.
Moreover, the preset traces $s$ need to be executed twice according to the strategies devised in 
\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}, because trace refinement and correctness of nondeterministic behaviour are checked by two different sets of test cases.

@@

\subsection{Overview}
In Section~\ref{section:preliminaries}, we present the background
relevant to our work. In Section~\ref{sec:finitecompletefails}, finite
complete test suites for verifying failures refinement are presented. 
A sample test
suite is presented in Section~\ref{sec:case}.
Test
suites checking trace refinement are a simplified version of the former
class; they are presented in Section~\ref{sec:finitecomplete}.
The optimality results are presented in Section~\ref{sec:complexity}, together with further
complexity considerations.
 Our results are discussed in
Section~\ref{sec:conc}, where we also conclude. References to further related work
are given throughout the paper where appropriate.

>>>>

\paragraph{Motivation}
Model-based testing (MBT) is an active research field; results are currently
being evaluated and integrated into industrial verification processes by many
companies worldwide. This holds particularly for the embedded and
cyber-physical systems domains, where critical systems require rigorous
testing~\cite{jp2018ets,DBLP:conf/isola/0001BH18}.

@@

Generation techniques for complete test suites have been developed for
various formalisms; we mention here representative work for finite state
machines~\cite{hierons_testing_2004,simao_reducing_2012}, timed
automata~\cite{Springintveld2001}, {\sf\em
Circus}~\cite{DBLP:journals/acta/CavalcantiG11}, Timed
CSP~\cite{Schneider:1995:OST:203471.203475}, general labelled transition
systems~\cite{DBLP:journals/cn/Tretmans96}, symbolic state
machines~\cite{DBLP:conf/icst/Petrenko16}, and Kripke
structures~\cite{Huang2017}. In this article, we tackle \emph{Communicating
Sequential Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a
mature process algebra that has been shown to be well-suited for the
description of reactive control systems in many publications over almost five
decades. Many of these applications are described in~\cite{Roscoe2010} and in
the references there. Industrial success has also been reported; see, for
example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.

@@

\paragraph{Main Contributions}

This article presents   complete black-box test suites for software and
systems modelled using CSP. They can be generated for non-terminating,
divergence-free, finite-state CSP processes with finite alphabets,
interpreted both in the trace and the failures semantics. Divergence freedom
is usually assumed in black-box testing, since it cannot distinguish between
divergence and deadlock using testing.

\newpage
\noindent %
The main novel contributions in this article may be summarised as
follows.
\begin{enumerate}
\item It is shown that trace or failures conformance can be established
    with finitely many test cases, provided suitable fault domains are
    chosen, so that the true behaviour of the SUT is reflected by members
    of these domains.

@@

\paragraph{Related Work}
Our results complement and extend work previously published
in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
None of these provide sufficient conditions for constructing finite complete
test suites. So, they also do not provide complexity bounds on the number of
test executions needed to establish conformance between an SUT and a
reference process. In~\cite{DBLP:conf/pts/CavalcantiS17}, fault domains are
used, but these contain all processes refining a ``top'' fault domain
process. This concept is orthogonal to the one investigated here:~members of
our fault domain need not be in refinement relation to any other process in
the domain. They just adhere to the same upper bound $q$ of nodes in their
normalised transition graphs.

@@

In~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17},
the authors devised linear test cases: after running through a preset trace
$s$, test cases for trace refinement check for illegal acceptance of a
specific event $e$, and test cases verifying  nondeterministic behaviour
check for  acceptance of events from some set $A$. In the present article, we
follow the alternative approach proposed
in~\cite{DBLP:conf/fm/PeleskaS96,peleska1997a} and use {\it adaptive} test
cases. This means that each test case adapts its trace execution to the
nondeterministic behaviour of the SUT, checks for trace violations at any
point during the test execution, and checks for the acceptance of a given
minimal hitting set of events after any legal trace of a test case-specific
length.

The adaptive test cases have the advantage that test executions only lead to
an inconclusive result if the reference process allows for a nondeterministic
choice between deadlock and trace continuation in a certain state. In
contrast to this, the linear test cases may lead to many more futile
executions with inconclusive results, if the SUT refuses to engage into the
next event $e$ from the preset trace $s$, due to legal nondeterministic
choices leading to a refusal of $e$. Moreover, the preset traces $s$ need to
be executed twice according to the strategies devised in
\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17},
because trace refinement and correctness of nondeterministic behaviour are
checked by two different sets of test cases.

@@

\paragraph{Overview} In Section~\ref{section:preliminaries}, we present the
background relevant to our work. In Section~\ref{sec:finitecompletefails},
finite complete test suites for verifying failures refinement are presented.
A sample test suite is presented in Section~\ref{sec:case}. Test suites
checking trace refinement are a simplified version of the former class; they
are presented in Section~\ref{sec:finitecomplete}. The optimality results are
presented in Section~\ref{sec:complexity}, together with further complexity
considerations. Our results are discussed in Section~\ref{sec:conc}, where we
also conclude. References to further related work are given throughout the
paper where appropriate.
